---
title: "ESM Personality Structure: Items Imputed"
author: "Alexander P. Christensen"
date: "5/31/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(EGAnet)
library(igraph)
library(NetworkToolbox)
library(corrplot)
library(knitr)
library(kableExtra)
```

```{r data, echo = FALSE, eval = TRUE, message = NA, warning = NA}
# Load data
## Imputation
### Directory
DIR <- "./02-data/04-item-wide/02-imputed/02-participant-data/csv/"

### Files
FILES <- list.files(DIR)

### Load
imp_items <- matrix(NA, nrow = 0, ncol = 62)

### Progress bar
pb <- txtProgressBar(max = length(FILES), style = 3)

for(i in 1:length(FILES)){
  
  # Items only
  items <- read.csv(paste(DIR, FILES[i], sep = ""))[,-c(61:62)]
  
  # Remove rows will all NA
  items <- items[-which(apply(items, 1, function(x){all(is.na(x))})),]
  
  # Add id and group for dynEGA
  items <- cbind(
    items,
    rep(gsub(".csv", "", FILES[i]), nrow(items)), # id
    rep(1, nrow(items)) # group
  )
  colnames(items)[61:62] <- c("ID", "Group")
  
  # Stack
  imp_items <- rbind(imp_items, items)
  
  # Update progress bar
  setTxtProgressBar(pb, i)
  
}

# Close progress bar
close(pb)

# Convert to data frame
imp_items <- as.data.frame(imp_items)

```

# Q1

## Dynamic Exploratory Graph Analysis: Population
### Traits
```{r dynEGA_population, echo = TRUE, eval = TRUE, message = NA, warning = NA}
imp_ega_pop <- dynEGA(
  data = imp_items,
  n.embed = 4,
  tau = 1, delta = 1,
  level = "population",
  id = ncol(imp_items) - 1, group = ncol(imp_items),
  use.derivatives = 1,
  model = "glasso", algorithm = "louvain",
  corr = "pearson",
  ncores = 6
)

plot(imp_ega_pop)
```

#Q2

## Dynamic Exploratory Graph Analysis: Population
### Facets
#### Openness to Experience
```{r dynEGA_openness, echo = TRUE, eval = TRUE, message = NA, warning = NA}
# Get openness items
imp_open <- imp_items[,c(
  grep("openness", colnames(imp_items)), 61, 62
)]

imp_ega_open <- dynEGA(
  data = imp_open,
  n.embed = 4,
  tau = 1, delta = 1,
  level = "population",
  id = ncol(imp_open) - 1, group = ncol(imp_open),
  use.derivatives = 1,
  model = "glasso", algorithm = "louvain",
  corr = "pearson",
  ncores = 6
)

plot(imp_ega_open)
```

#### Conscientiousness
```{r dynEGA_conscientiousness, echo = TRUE, eval = TRUE, message = NA, warning = NA}
# Get conscientiousness items
imp_cons <- imp_items[,c(
  grep("conscientiousness", colnames(imp_items)), 61, 62
)]

imp_ega_cons <- dynEGA(
  data = imp_cons,
  n.embed = 4,
  tau = 1, delta = 1,
  level = "population",
  id = ncol(imp_cons) - 1, group = ncol(imp_cons),
  use.derivatives = 1,
  model = "glasso", algorithm = "louvain",
  corr = "pearson",
  ncores = 6
)

plot(imp_ega_cons)
```

#### Extraversion

```{r dynEGA_extraversion, echo = TRUE, eval = TRUE, message = NA, warning = NA}
# Get extraversion items
imp_extr <- imp_items[,c(
  grep("extraversion", colnames(imp_items)), 61, 62
)]

imp_ega_extr <- dynEGA(
  data = imp_extr,
  n.embed = 4,
  tau = 1, delta = 1,
  level = "population",
  id = ncol(extr) - 1, group = ncol(extr),
  use.derivatives = 1,
  model = "glasso", algorithm = "louvain",
  corr = "pearson",
  ncores = 6
)

plot(imp_ega_extr)
```

#### Agreeableness

```{r dynEGA_agreeableness, echo = TRUE, eval = TRUE, message = NA, warning = NA}
# Get agreeableness items
imp_agre <- imp_items[,c(
  grep("agreeableness", colnames(imp_items)), 61, 62
)]

imp_ega_agre <- dynEGA(
  data = imp_agre,
  n.embed = 4,
  tau = 1, delta = 1,
  level = "population",
  id = ncol(agre) - 1, group = ncol(agre),
  use.derivatives = 1,
  model = "glasso", algorithm = "louvain",
  corr = "pearson",
  ncores = 6
)

plot(imp_ega_agre)
```

#### Neuroticism

```{r dynEGA_neuroticism, echo = TRUE, eval = TRUE, message = NA, warning = NA}
# Get neuroticism items
imp_neur <- imp_items[,c(
  grep("neuroticism", colnames(imp_items)), 61, 62
)]

imp_ega_neur <- dynEGA(
  data = imp_neur,
  n.embed = 4,
  tau = 1, delta = 1,
  level = "population",
  id = ncol(neur) - 1, group = ncol(neur),
  use.derivatives = 1,
  model = "glasso", algorithm = "louvain",
  corr = "pearson",
  ncores = 6
)

plot(imp_ega_neur)
```

## Loading Comparison Test
### Traits

```{r LCT_traits, echo = TRUE, eval = TRUE, message = NA, warning = NA}
LCT(imp_ega_pop$dynEGA$cor.data, n = length(FILES))
```

### Facets
#### Openness to Experience
```{r LCT_open, echo = TRUE, eval = TRUE, message = NA, warning = NA}
LCT(imp_ega_open$dynEGA$cor.data, n = length(FILES))
```

#### Conscientiousness
```{r LCT_cons, echo = TRUE, eval = TRUE, message = NA, warning = NA}
LCT(imp_ega_cons$dynEGA$cor.data, n = length(FILES))
```

#### Extraversion
```{r LCT_extr, echo = TRUE, eval = TRUE, message = NA, warning = NA}
LCT(imp_ega_extr$dynEGA$cor.data, n = length(FILES))
```

#### Agreeableness
```{r LCT_agre, echo = TRUE, eval = TRUE, message = NA, warning = NA}
LCT(imp_ega_agre$dynEGA$cor.data, n = length(FILES))
```

#### Neuroticism
```{r LCT_neur, echo = TRUE, eval = TRUE, message = NA, warning = NA}
LCT(imp_ega_neur$dynEGA$cor.data, n = length(FILES))
```

# Q3

## Dynamic Exploratory Graph Analysis: Individual

```{r dynEGA_individual, echo = TRUE, eval = TRUE, message = NA, warning = NA}
imp_ega_ind <- dynEGA(
  data = imp_items,
  n.embed = 4,
  tau = 1, delta = 1,
  level = "individual",
  id = ncol(imp_items) - 1, group = ncol(imp_items),
  use.derivatives = 1,
  model = "glasso", algorithm = "louvain",
  corr = "pearson",
  ncores = 6
)
```

# Q4

## Distribution of Within-person Factors
```{r dynEGA_individual_distribution, echo = TRUE, eval = TRUE, message = NA, warning = NA}
# Obtain dimensions
dims <- unlist(
  lapply(imp_ega_ind$dynEGA, function(x){
  x$n.dim
})
)

# Summary of dimensions
desc.dims <- c(
  mean(dims), sd(dims), range(dims)
)
names(desc.dims) <- c("Mean", "SD", "Low", "High")

# Proportions
dimensions <- list()

dimensions$Between <- ega_pop$dynEGA$n.dim

dimensions$Within <- table(unlist(lapply(ega_ind$dynEGA, function(x){x$n.dim}))) / length(ega_ind)
```

## Match between Theoretical Structure
```{r NMI_theoretical, echo = TRUE, eval = TRUE, message = NA, warning = NA}
# Obtain item placements
placements <- lapply(imp_ega_ind$dynEGA, function(x){
  x$wc
})

# Theoretical placements
theoretical <- numeric(length = ncol(imp_items[,-c(61,62)]))
theoretical[c(grep("openness", colnames(imp_items)))] <- 1
theoretical[c(grep("conscientiousness", colnames(imp_items)))] <- 2
theoretical[c(grep("extraversion", colnames(imp_items)))] <- 3
theoretical[c(grep("agreeableness", colnames(imp_items)))] <- 4
theoretical[c(grep("neuroticism", colnames(imp_items)))] <- 1
names(theoretical) <- names(placements[[1]])


# Loop through placements
theo_nmi <- unlist(
  lapply(
  placements, function(x, theo){
    
    # Remove items not in a dimension
    items <- na.omit(x)
    
    # Match with theoretical
    matched <- theo[match(names(items), names(theo))]
    
    # Compute NMI
    igraph::compare(items, matched, method = "nmi")
    
  }, theo = theoretical
)
)

```

## Match between People
```{r NMI_people, echo = TRUE, eval = TRUE, message = NA, warning = NA}

# Initialize matrix of people
people_nmi <- matrix(nrow = length(FILES), ncol = length(FILES))
colnames(people_nmi) <- gsub(".csv", "", FILES)
row.names(people_nmi) <- colnames(people_nmi)

# Loop through placements
for(i in 1:length(FILES))
  for(j in 1:length(FILES)){
    
    # Person 1
    person1 <- na.omit(placements[[paste(
      "ID", gsub(".csv", "", FILES[i]), sep = ""
    )]])
    
    # Person 2
    person2 <- na.omit(placements[[paste(
      "ID", gsub(".csv", "", FILES[j]), sep = ""
    )]])
    
    # Match person 1 with person 2
    matched <- intersect(names(person1), names(person2))
    
    # Compute NMI
    people_nmi[i,j] <- igraph::compare(
      person1[matched], person2[matched], method = "nmi"
    )
    
  }

# Visualize
col0 <- colorRampPalette(c("grey", "orange", "#FF7733"))

corrplot(
  people_nmi, method = "color",
  cl.lim = c(0,1), is.corr = FALSE,
  type = "lower", diag = FALSE,
  addgrid.col = "black",
  tl.col = "black", col = col0(100),
  na.label.col = "white"
)
```

#Q7

## Rank-order Correlation
```{r centrality, echo = TRUE, eval = TRUE, message = NA, warning = NA}
# Initialize matrix for strengths
strengths <- matrix(nrow = ncol(imp_ega_pop$dynEGA$network), ncol = length(FILES) + 1)
colnames(strengths) <- c("Between", gsub(".csv", "", FILES))
row.names(strengths) <- colnames(imp_ega_pop$dynEGA$network)

# Loop through placements
for(i in 1:(length(FILES) + 1)){
  
  if(i == 1){
    strengths[,i] <- strength(imp_ega_pop$dynEGA$network)
  }else{
    strengths[,i] <- strength(imp_ega_ind$dynEGA[[(i-1)]]$network)
  }
  
}

# Rank-order correlation with between-person strength
cor(strengths, method = "spearman", use = "pairwise.complete.obs")[-1,1]
```

## How many nodes in the top five of between-person are the same as top five within-person

```{r top_five, echo = TRUE, eval = TRUE, message = NA, warning = NA}
# Get rank-orderings
strength_rank <- apply(strengths, 2, rank)

# Make list of top five
top_fives <- apply(strength_rank, 2, function(x){
  as.list(x[x <= 5])
})

# Obtain the matches across people
matched_top_fives <- lapply(top_fives[-1], function(x, between){
  na.omit(match(names(x), names(between)))
}, between = top_fives[[1]])

# Compute proportions
top_five_proportions <- table(unlist(matched_top_fives)) / length(matched_top_fives)

# Manage table structure
names(top_five_proportions) <- names(top_fives[[1]][as.numeric(names(top_five_proportions))])
names(top_five_proportions) <- gsub(".Ord1", "", names(top_five_proportions))
top_five_proportions <- c(top_five_proportions, length(top_five_proportions) / ncol(imp_ega_pop$dynEGA$network))
names(top_five_proportions)[length(top_five_proportions)] <- "Random Chance"
top_five_proportions <- as.data.frame(top_five_proportions)
colnames(top_five_proportions) <- "Proportion"

# Create table
kable(
  top_five_proportions, row.names = TRUE
)

```


